### 2022-04-26 LaTeX_Plos/ResponseToReviewers.tex

\date{\today}

\begin{document}

\begin{letter}{}

\opening{Dear Dr.~Annamalai:}

We would like to thank each of the five reviewers for their evaluation of our manuscript and generally favorable reviews.  
%which has helped us revise the manuscript.  %Both reviewers mention previous research that has been performed on  silver nanoparticles in textiles.  It should be emphasized that our research focuses on reactive silver inks for textiles and their improved antiviral performance and wash durability compared to silver nanoparticles.
%Our work focuses on antiviral activity after washing, which has not been previously reported for silver nanoparticles and ink.  
%Additionally, in our submitted paper, we test the fabric treatments after extended bleach washing with a harsh ultrasonication washing procedure under high concentration bleach (10\%) which has not been previously reported. 
%to allow our work to be of interest 
%to the readership of ACS Applied Nano Materials.
% We thank the three reviewers for their evaluation of our manuscript and comments.  
%We thank the reviewers' statements that ``this review article is comprehensive to introduce the various sources of bio-inspiration, the critical performance requirements and the strategies that have been carried out to tackle the challenges" and we have provided ``an elaborate and comprehensive arrangement of various transparent substrates and barrier layers for optoelectronics." 
We have made improvements to our paper based on the reviewers' comments and suggestions.
Below, we summarize the changes that have been made and respond to the reviewers' comments.
Considering the importance of the topic of this
article on 
%for application of 
bleach washable, anti-viral fabric coatings,
we resubmit this manuscript for your consideration of publication in PLOS One.
Thank you for your consideration.

@@

\blue{Citations of these recent works have been added to the text.}

\textbf{Reviewer: 1}

\textbf{Recommendation:} This paper investigates the use of silver nanoparticle and inks incorporated into knitted medical textiles that show durable anti-viral activity, even after extended washing in bleach. 
The paper is interesting, generally well-written and organised. 
However, several issues need to be addressed before it can be considered for publication. Thus, I would recommend major changes to the manuscript before resubmission and reconsideration.

\blue{We thank the reviewer for their assessment that our paper is ``interesting, generally well-written and organised."  The manuscript has been updated based on their comments below.}

@@

\textbf{Comments:} 
1. Uniform distribution of silver on the PET fabrics is crucial to the good antiviral property. How did the authors adjust silver distribution via drop casting method? How about the silver content on the coating? Cytotoxicity is needed to characterize the coated PET fabrics.

\blue{PET samples were dampened with alcohol to improve the spreading of the silver coatings during drop casting. The silver content of the coating has been added in the Materials and Methods section. Cytotoxicity assessment is outside the scope of the work.  Line 253 in the Conclusion has been updated to state, ``Future work should analyze the antiviral mechanism, as well as the cytotoxicity of reactive silver ink treated fabrics before
commercial use."}

2. The authors need to characterize PDMS coatings, for example, the thickness of PDMS coating? How did reactive silver ink and silver nanoparticle influence the final morphology of PDMS coatings?

\blue{A description of the final morphology of samples after PDMS coatings have been added in page 4, paragraph 4 line 150 ,``The PDMS
coating appears roughened due to the underlying morphology of the silver nanoparticles
or silver ink coated fabrics." }
\commentPaul{Anything about thickness?}
\commentAnthony{I do not have exact measurements. I have some pictures but they are low quality. We could say less than 1 micron. }
\commentPaul{Yes, perhaps, say less than about 1 micron  in the text.}

@@

\blue{The PDMS coating adds a repellency property to the fabric, so there is simultaneous virus deactivation and repellency. ``The PDMS coated silver 
fabrics may release less silver than without PDMS; however, a greater reduction of 
infectious virus is observed with PDMS from a combination of virus inactivation and 
liquid repellency" has been added to page 6, bottom of paragraph 2, line 224-226.}

\newpage
\textbf{Reviewer: 3}

\textbf{Recommendation:} This is a well-written article covering a nice piece of timely research, at a moment in history when there is considerable interest in antiviral textiles. The quality of the science is high and the conclusions from the article are generally very robust. However, I have a small number of concerns that should be addressed through minor revision of the MS, mainly in highlighting the limitations of the study. i.e. One might be forgiven from interpreting the work and resulting technology as being at a late stage of technical readiness so the following limitations should be transparently disclosed in the article:

\blue{We thank the Reviewer for their very positive assessment of the importance of our work as well as their positive assessment of the quality of the science and conclusions in our work.  }

@@

\textbf{Comments:} 
1. The efficacy of the reactive silver approach is clear, but are there any significant unknown risks such as toxicity or has the oxidation to elemental silver been proven to be 100\% reliable?

\blue{The ink leaves a pure silver film, so toxicity should not be different than elemental silver. The curing temperature is low and at this temperature, the silver is not oxidised during curing. The XRD characterization in Figure S2 shows the silver is not oxidized on the fabric after curing. \commentPaul{Perhaps add a statement in the text clarifying this point as well.}}

@@

4. Other properties - technical textiles don't just need to exhibit their desired primary benefit (in this case antiviral); does the treatment compromise any other aspects of textile properties that could be relevant to its use such as flexibility and comfort in wear?

\blue{We appreciate the suggestion and are considering analysis of flexibility and comfort 
in future work.}
%\blue{Future work may consider other textile properties and this limitation is acknowledged.  }

Finally, the descriptions of materials in the Materials and Methods are too vague and this prevents the work from being easily duplicated. Trademarks such as Antocon(R) are not acknowledged which is a product brand name not a supplier. That range of wipes contains more than one PET knit wipe so the specific one used should be described. The specific bleach is not given, only that it was purchased from Giant Eagle - was it Clorox(R) or their own brand? Do they really only sell one? Could product codes be provided for other purchases from Sigma-Aldrich unless they really only sell one PBS, FBS and PDMS product?

\blue{The supplier has been added, as well as a better description of the materials and methods. The specific bleach is Giant Eagle brand and ``(Giant Eagle brand, SKU: 00030034936150)" has been added to the text. 
%\commentPaul{Perhaps better description of this bleach.  Concentration level.  Sodium hypochlorite concentration level}
A better description of the other materials that were used have been changed in the Materials and Methods section. Product numbers for PBS (P5493) and FBS (12103C) has been added to the text. PDMS (Sylgard 184) has been added to the text. 
%\commentPaul{Try to provide better description of materials even if there is only one.  Sometimes, companies will change how many they sell or change the composition of the one product they sell.}
}

@@

*Both PET-NP and PET-Ink samples are dipped in PDMS (1:10 ratio, Sylgard) and cured in oven ... $-->$ What does 1:10 ratio stand for? Is it the PDMS: Hexane ratio or is it PET-NP or PET-Ink: PDMS ratio? This part is not clear. Authors should clarify this part. Fig 3a gives the schematic representation of PDMS treatment for silver coated PET samples and as it is understood from the figure PDMS is dissolved in Hexane however, authors did not mention about hexane in the article. The amounts and preparation steps of PDMS solution should be clarified. Also, Sylgard has various silicone elastomers with different codes. It is better to give the SylgardTM and code.

\blue{Sylgard 184 and a 1:10 ratio of curing agent:PDMS dissolved in hexane has been used. The description of PDMS has been modified within the text for clarity. }

*It would be great if the unwashed and washed fabrics antiviral activity tests were done simultaneously to keep the initial viral concentration same for all samples to see the difference between wash cycles. According to the data authors reveal for sample PET-NP/P antiviral activity is increased even a little with washing which seems not possible.

\blue{The observed change is not statistically significant.}
%is is not statistically true, even though the overall mean may be slightly less. The groups are not statistically different.}

@@

* For XRD peaks are observed between 20-50. Ag has two more major peaks at 64.7 and 77.4. It would be better to scan till 80$\degree$.

\blue{The authors choose to show the largest peaks in the 20-50$\degree$ range. }

@@

\blue{We thank the reviewer for acknowledging the importance of our study.}

\textbf{Comments:}

53 In this study the silver amine complex is being studied, however, this ink needs to be fully chemically characterised and formulated.

\blue{The silver amine complex has been fully chemically characterised and formulated in the cited paper 39.}  %Additional formulation is held under intellectual property.}

@@

101 The x-axis of the XRD pattern should change to (2$\theta$ (°)). Please at Y-axis please remove numbers if it is going to be reported by Arbitrary unit or remove (a.u.).

\blue{The axes have been changed. }

104 Author mentioned that the untreated PET is fully wetted by water and Fig 2 (a) shows that the water static, contact angle, for PET is zero. Although other studies such as the Indian Journal of Fibre \& Textile Research Vol. 37, September 2012, pp. 287-291 showed the water contact angle of untreated PET fabric is 124.24$\degree$.

\blue{The paper mentioned uses weaved PET fabric which may be more repellent. The PET we used is knit and is fabricated to have high sorbency. }
%\commentPaul{See above comment about PET having lower surface energy than other textiles. Perhaps, this PET is not so low surface energy?}  
%\commentAnthony{Maybe. I think the reviewer just forgot to add with PDMS to make a low surface energy surface. We never say PET itself is low surface energy.}

### 2022-04-26 LaTeX_Plos/ResponseToReviewers.tex

\textbf{Comments:} 
1. Uniform distribution of silver on the PET fabrics is crucial to the good antiviral property. How did the authors adjust silver distribution via drop casting method? How about the silver content on the coating? Cytotoxicity is needed to characterize the coated PET fabrics.

\blue{PET samples were dampened with alcohol to improve the spreading of the silver coatings during drop casting. The silver content of the coating has been added in the Materials and Methods section. Cytotoxicity assessment is outside the scope of the work.  Line 253 in the Conclusion has been updated to state, ``Future work should analyze the antiviral mechanism, as well as the cytotoxicity of reactive silver ink treated fabrics before
commercial use."}

2. The authors need to characterize PDMS coatings, for example, the thickness of PDMS coating? How did reactive silver ink and silver nanoparticle influence the final morphology of PDMS coatings?

\blue{A description of the final morphology of samples after PDMS coatings have been added in page 4, paragraph 4 line 150 ,``The PDMS layer is observed to be less than 1 micron in thickness from SEM. The PDMS
coating appears roughened due to the underlying morphology of the silver nanoparticles
or silver ink coated fabrics." }
%\commentPaul{Anything about thickness?}
%\commentAnthony{I do not have exact measurements. I have some pictures but they are low quality. We could say less than 1 micron. }
%\commentPaul{Yes, perhaps, say less than about 1 micron  in the text.}

@@

\textbf{Recommendation:} This is a well-written article covering a nice piece of timely research, at a moment in history when there is considerable interest in antiviral textiles. The quality of the science is high and the conclusions from the article are generally very robust. However, I have a small number of concerns that should be addressed through minor revision of the MS, mainly in highlighting the limitations of the study. i.e. One might be forgiven from interpreting the work and resulting technology as being at a late stage of technical readiness so the following limitations should be transparently disclosed in the article:

\blue{We thank the Reviewer for their very positive assessment of the importance of our work as well as their positive assessment of the quality of the science and conclusions in our work.  }

\textbf{Comments:} 
1. The efficacy of the reactive silver approach is clear, but are there any significant unknown risks such as toxicity or has the oxidation to elemental silver been proven to be 100\% reliable?

\blue{The ink leaves a pure silver film, so toxicity should not be different than elemental silver. The curing temperature is low and at this temperature, the silver is not oxidised during curing. "The XRD characterization also shows the silver is not oxidized on the fabric after curing" has been added to the text on page 3, line 102-103. %\commentPaul{Perhaps add a statement in the text clarifying this point as well.}
}
